Adapting Duplicate Filters to a Sliding Window Context

### 2020-03-20 qhtv2.tex
Améliorations variées

	When analysed in the limit of an infinite stream, error rates of duplicate filters turn out to be heavily constrained; as a consequence they appear to provide no advantage, asymptotically, over a random yes-no coin toss \cite{10.1145/3297280.3297335}.

@@

Finally, it is clear that the input distribution plays a central role regarding how efficiently the DDP can be solved. For instance, some deterministic streams may be expressed very compactly (such as the output of a PRNG with known seed) making the DDP relatively easy. Information-theoretically, if the source has $U$ bits of entropy then the situation is equivalent to having an $U$-bit, uniformly distributed input.\footnote{Another motivation for this is pragmatic: in many applications, duplicate detection is performed on \emph{hashes} of actual data rather than on the data itself, and we may assume that hashes are uniformly distributed.}

\todo[continuer].

\subsection{Contributions}
We introduce a variant of the DDP over a sliding window, which enables the design of filters that avoids the saturation issue identified in \cite{10.1145/3297280.3297335} (Section~\ref{sec:filsat}). We provide explicit upper bounds, and corresponding constructions, for the amount of memory necessary to solve this problem (Section~\ref{sub:opt_solutions}), then introduce the \emph{queuing construction} as a generic approach to build windowed-DDP-friendly filters from existing filters (Sections~\ref{sec:queue} and \ref{sec:subfilter-select}). We analyse the queuing construction's error rates and parameters (Section~\ref{sec:optimising}), which are confirmed by benchmarks (Section~\ref{sec:bench}). Finally, we provide an analysis of the queueing construction's resistance to adversarial streams (Section~\ref{sec:adversarial}).

\subsection{Related work}\label{sub:related}
\todo[continuer]
In \cite{10.1145/1060745.1060753}, the authors explore the behaviour of a structure very close to the counting Bloom filter in some specific sliding window context, that they call landmark- and jumping sliding windows. However, they do not deal with the problematic of a sliding window where at each new element added to the sliding window, another element is removed.
Closer to our work, the idea of using subfilters, tossing old ones has already been explored in the A2 filter \cite{Yoo10}. However, the author only considered a small portion of the problem, as they only discussed about using two subfilters, and only Bloom Filters acting as subfilters.

In this work, we provide a full abstraction of the mechanism, allowing us to use many other DDF than Bloom Filters. We also provide a generalised analysis of the error rates.
\subsection{Organisation}

\section{Duplicate Detection and Filter Saturation}
\subsection{Notations and Initial Problem Statement}
Let us consider an alphabet $\Gamma$. A stream $E$ over $\Gamma$ is a (possibly infinite) set $E = (e_1, e_2, \dotsc)$, with each $e_i \in \Gamma$.

@@

\begin{proof}
	Here again we explicitely construct the filters that achieve the theorem's bounds.

@@

\subsection{The queueing construction}

@@

When inserting a new element in the queueing filter, it is inserted in the topmost subfilter of the queue. After $c$ insertions, a new empty filter is added to the queue, and the oldest subfilter is popped and erased.

@@

\paragraph{Pseudocode.}
We give a brief pseudo-code of the queuing filter's functions \textsf{Lookup} and \textsf{Insert}, as well as a \textsf{Setup} function for initialisation, in Algorithm~\ref{alg:queueing}. We introduced for simplicity a constructor $\mathcal F.\textsf{Setup}$ that takes as input an integer $M$ and outputs an initialized empty filter $\mathcal F$ of size at most $M$. The \texttt{subfilters} is a FIFO that has a \texttt{pop} and \texttt{push\_first} operations, which respectevely removes the last element in the queue or inserts a new item in first  position.

\begin{algorithm}[]
	\caption{Queueing Filter Setup, Lookup and Insert}\label{alg:queueing}

@@

\subsection{Error rate analysis}
The queueing filter's properties can be derived from the subfilters'. False positive and false negative rates are of particular interest. In this section we consider a queuing filter $\mathcal Q$ with $L$ subfilters of type $\mathcal F$ and 
capacity $c$ (which means that the last subfilter is dropped after $c$ insertions).

@@

\section{Optimising queuing filters}
\label{sec:optimising}
\todo[À vérifier je suis pas convaincu encore]
\begin{theorem}
	In first approximation, a low value of $L$ for constant value $cL$ leads to a higher false negative probability for the queueing filter.
\end{theorem}

\begin{proof}
As a matter of fact, on average the queueing filter consists of $L-1$ subfilters with $c$ insertions, and one subfilter with $\frac c2$ insertions. As such, we can say that the queueing filter is a representation of the last $(c-\frac12) L$ last elements.

@@

\section{Queuing filters from existing DDFs}\label{sec:subfilter-select}
Our queuing construction relies on a choice of subfilters. A first observation 
is that we may assume that all subfilters can be instances of a single 
DDF design (rather than a combination of different designs).

@@

It remains to decide which subfilter construction to choose.

\subsection{Subfilter selection}
\label{sec:optimal_ddf}
When comparing different DDFs, one interesting metric in our context is how fast
such filters saturate: indeed, in the original DDP setting, all filters eventually saturate \cite{10.1145/3297280.3297335}.

@@

In the experiments we use QHT for the benchmarking of the queueing construction.

\subsection{How many subfilters?}
The parameter $L$ determines how many subfilters appear in the queueing construction. Summing up the false positive and false negative rates, we have a total error rate 
\begin{equation}
	\operatorname{ER}_{\mathcal Q, cn}^{w}
	= 1 - \alpha \beta^{L-1} + \alpha' \beta'^{L-1}
\end{equation}
where $\beta = 1 - \FP_{\mathcal F, c}$, $\beta' = u_c$, 
$\alpha = \frac1c \sum_{\ell = 0}^{c-1}1 - \FP_{\mathcal F, \ell}$ and $\alpha' = \frac1c \sum_{\ell = 0}^{c-1} u_\ell$ only depend on $c$ and the choice of subfilter type $\mathcal F$. Differentiating with respect to $L$ and equating to zero gives an optimum
\begin{equation*}
	L = -\log((\alpha \beta' \log(\beta))/(\alpha'\beta \log(\beta')))/\log(\beta/\beta').
\end{equation*}

@@

\subsection{Benchmarking Queuing Filters}
We now use the queuing construction on the DDFs discussed previously. In Section~\ref{sec:optimal_ddf} we suggested the heuristic that the DDFs with the least saturation rate in the DDP would yield the best (error-wise) queueing filter for the sliding window DDP. This heuristic is supported by results, summarized in Figure~\ref{tab:comparing_queueing}. For this benchmark we used the following parameters: uniform stream from an alphabet of size $|\Gamma| = 2^{18}$, memory size $M=100,000$ bits, sliding window of size $w = 10,000$, and we measure the error rate (sum of $\FNR^w$ and $\FPR^w$).

@@

\begin{figure}[!ht]
	\input{graphs/compare_queues.tex}
	\caption{Error rate (times 100) of various queueing filters on various sliding window sizes, $M=100000$, $L=10$, $|\Gamma| = 2^{18}$.}\label{tab:comparing_queueing}
\end{figure}
%\begin{table}[]
%	\begin{tabular}{rccccc}
%			   & $w=1000$ & $w=5000$ & $w=10^4$ & $w=10^5$ & $w=10^6$ \\ \hline
%		QHT    & \textbf{9.84} 	  & \textbf{26.25}	 & \textbf{41.09}	 & \textbf{87.89}	  & 70.76      \\
%		Cuckoo & 84.99	  & 99.94	 & 99.95	 & 99.93	  & 99.80      \\
%		SQF    & 16.73 	  & 48.76	 & 69.25	 & 95.77 	  & \textbf{70.14}     \\
%		SBF    & 54.88 	  & 93.93	 & 97.33	 & 97.41	  & 71.17      \\
%	\end{tabular}
%	\caption{Error rate (times 100) of various queueing filters on various sliding window sizes, $M=100000$, $L=10$, $|\Gamma| = 2^{16}$. For each sliding window, the best performance is written in bold.}\label{tab:comparing_queueing}
%\end{table}

\subsection{How many subfilters?}\label{sub:L_bench}
\todo[À faire]

\subsection{Comparison of QHT and QHT-Queueing Filter}
\todo[À faire]
In this section, we compare the efficiency of QHT and QHT-queueing filter. For this, we use both an artificial dataset, as well as a real one, on various sliding window sizes.

>>>>

	When analysed in the limit of an infinite stream, error rates of duplicate filters turn out to be heavily constrained; as a consequence they appear to provide no advantage, asymptotically, over a random biased coin toss \cite{10.1145/3297280.3297335}.

@@

On a side note, it is clear that the input distribution plays a central role regarding how efficiently the DDP can be solved. For instance, some deterministic streams may be expressed very compactly (such as the output of a PRNG with known seed) making the DDP relatively easy. Information-theoretically, if the source has $U$ bits of entropy then the situation is equivalent to having an $U$-bit, uniformly distributed input.

Several algorithms have been proposed to solve the DDP, but
they \emph{all} face a phenomenon called \enquote{saturation} on large data streams \cite{10.1145/3297280.3297335}. Essentially, when saturation happens the algorithm performs no better than at random. This is problematic on two grounds: it makes the comparison of several algorithms difficult (since they all asymptotically behave in that fashion), and the unavoidable saturation ruins any particular design's merits.

\subsection{Contributions}
To circumvent the aforementioned limitations, we introduce a variant of the DDP \emph{over a sliding window}, which enables the design of filters that avoids the saturation issue altogether (Section~\ref{sec:dudefisa}). We provide explicit upper bounds for the amount of memory necessary to solve this problem, and corresponding constructions.

We then introduce the \emph{queuing construction}, a generic approach to build windowed-DDP-friendly filters from existing filters (Sections~\ref{sec:queue} and \ref{sec:subfilter-select}). We analyse the queuing construction's error rates and parameters (Section~\ref{sec:optimising}), which are confirmed by benchmarks (Section~\ref{sec:bench}).

Finally, we provide an analysis of the queuing construction's resistance to adversarial streams (Section~\ref{sec:adversarial}).

\subsection{Related work}\label{sub:related}
\cite{10.1145/1060745.1060753} studies a structure aking to the counting Bloom filter, in a \enquote{landmark-} and \enquote{jumping-} sliding window setting. Despite the name, these are different from the sliding window discussed in this paper.

Closer to our work, the idea of using subfilters may be found in the A2 filter's design \cite{Yoo10}. The A2 is built from two Bloom filters, a situation which we generalise and analyse generically in this paper.

\todo[continuer]

\subsection{Organisation}

\section{Duplicate Detection and Filter Saturation}
\label{sec:dudefisa}
\subsection{Notations and Initial Problem Statement}
Let us consider an alphabet $\Gamma$. A stream $E$ over $\Gamma$ is a (possibly infinite) set $E = (e_1, e_2, \dotsc)$, with each $e_i \in \Gamma$.

@@

\begin{proof}
	Here again we explicitly construct the filters that attain the theorem's bounds.

@@

\subsection{The queuing construction}

@@

When inserting a new element in the queuing filter, it is inserted in the topmost subfilter of the queue. After $c$ insertions, a new empty filter is added to the queue, and the oldest subfilter is popped and erased.

@@

\paragraph{Pseudocode.}
We give a brief pseudocode for the queuing filter's functions \textsf{Lookup} and \textsf{Insert}, as well as a \textsf{Setup} function for initialisation, in Algorithm~\ref{alg:queuing}. We introduced for simplicity a constructor $\mathcal F.\textsf{Setup}$ that takes as input an integer $M$ and outputs an initialized empty filter $\mathcal F$ of size at most $M$. Here \texttt{subfilters} is a FIFO that has a \texttt{pop} and \texttt{push\_first} operations, which respectevely removes the last element in the queue or inserts a new item in first position.\footnote{Two standard ways to implement this are a double-ended queue (deque) and a ring buffer.}

\begin{algorithm}[]
	\caption{Queuing Filter Setup, Lookup and Insert}\label{alg:queuing}

@@

\subsection{Error rate analysis}
The queuing filter's properties can be derived from the subfilters'. False positive and false negative rates are of particular interest. In this section we consider a queuing filter $\mathcal Q$ with $L$ subfilters of type $\mathcal F$ and 
capacity $c$ (which means that the last subfilter is dropped after $c$ insertions).

@@

\section{Optimising queuing filters}
\label{sec:optimising}

The parameter $L$ determines how many subfilters appear in the queuing construction. Summing up the false positive and false negative rates, we have a total error rate 
\begin{equation}
	\operatorname{ER}_{\mathcal Q, cn}^{w}
	= 1 - \alpha \beta^{L-1} + \alpha' \beta'^{L-1}
\end{equation}
where $\beta = 1 - \FP_{\mathcal F, c}$, $\beta' = u_c$, 
$\alpha = \frac1c \sum_{\ell = 0}^{c-1}1 - \FP_{\mathcal F, \ell}$ and $\alpha' = \frac1c \sum_{\ell = 0}^{c-1} u_\ell$ only depend on $c$ and the choice of subfilter type $\mathcal F$. Differentiating with respect to $L$ and equating to zero gives an optimum
\begin{equation*}
	L = -\log((\alpha \beta' \log(\beta))/(\alpha'\beta \log(\beta')))/\log(\beta/\beta').
\end{equation*}

\todo[À vérifier la suite je suis pas convaincu encore]
\begin{theorem}
	In first approximation, a low value of $L$ for constant value $cL$ leads to a higher false negative probability for the queuing filter.
\end{theorem}

\begin{proof}
As a matter of fact, on average the queuing filter consists of $L-1$ subfilters with $c$ insertions, and one subfilter with $\frac c2$ insertions. As such, we can say that the queuing filter is a representation of the last $(c-\frac12) L$ last elements.

@@

\section{Queuing filters from existing DDFs}
\label{sec:subfilter-select}
\label{sec:optimal_ddf}
Our queuing construction relies on a choice of subfilters. A first observation 
is that we may assume that all subfilters can be instances of a single 
DDF design (rather than a combination of different designs).

@@

It remains to decide which subfilter construction to choose. 
When comparing different DDFs, one interesting metric in our context is how fast
such filters saturate: indeed, in the original DDP setting, all filters eventually saturate \cite{10.1145/3297280.3297335}, but they may reach that state more or less quickly. The following result gives the behaviour of a perfect filter regarding this question:

@@

%In the experiments we use QHT for the benchmarking of the queuing construction.

@@

\subsection{Benchmarking Queuing Filters}
We now use the queuing construction on the DDFs discussed previously. In Section~\ref{sec:optimal_ddf} we suggested the heuristic that the DDFs with the least saturation rate in the DDP would yield the best (error-wise) queuing filter for the sliding window DDP. This heuristic is supported by results, summarized in Figure~\ref{tab:comparing_queuing}. For this benchmark we used the following parameters: uniform stream from an alphabet of size $|\Gamma| = 2^{18}$, memory size $M=100,000$ bits, sliding window of size $w = 10,000$, and we measure the error rate (sum of $\FNR^w$ and $\FPR^w$).

@@

\begin{figure}[!ht]
	\input{graphs/compare_queues.tex}
	\caption{Error rate (times 100) of various queuing filters on various sliding window sizes, $M=100000$, $L=10$, $|\Gamma| = 2^{18}$.}\label{tab:comparing_queuing}
\end{figure}
%\begin{table}[]
%	\begin{tabular}{rccccc}
%			   & $w=1000$ & $w=5000$ & $w=10^4$ & $w=10^5$ & $w=10^6$ \\ \hline
%		QHT    & \textbf{9.84} 	  & \textbf{26.25}	 & \textbf{41.09}	 & \textbf{87.89}	  & 70.76      \\
%		Cuckoo & 84.99	  & 99.94	 & 99.95	 & 99.93	  & 99.80      \\
%		SQF    & 16.73 	  & 48.76	 & 69.25	 & 95.77 	  & \textbf{70.14}     \\
%		SBF    & 54.88 	  & 93.93	 & 97.33	 & 97.41	  & 71.17      \\
%	\end{tabular}
%	\caption{Error rate (times 100) of various queuing filters on various sliding window sizes, $M=100000$, $L=10$, $|\Gamma| = 2^{16}$. For each sliding window, the best performance is written in bold.}\label{tab:comparing_queuing}
%\end{table}

\subsection{The number of subfilters}\label{sub:L_bench}
\todo[À faire]

\subsection{Comparison of QHT and QHT-Queuing Filter}
\todo[À faire]
In this section, we compare the efficiency of QHT and QHT-queuing filter. For this, we use both an artificial dataset, as well as a real one, on various sliding window sizes.
